---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "images/README-"
)

```

# GEKS-Törnqvist Decomposition


This package focuses on one function `GEKSdecomp`, which decomposes the multilateral GEKS-Tornqvist index into the contributions of individual commodities. The calculation is based purely on the *Decomposing Multilateral Price Indexes into the Contributions of Individual Commodities* paper written by *Michael Webster* and *Rory C. Tarnow-Mordi*. It uses the techinque 
described under section [2.2 Decomposition of the CCD and GEKS Methods](https://www.researchgate.net/publication/333875150_Decomposing_Multilateral_Price_Indexes_into_the_Contributions_of_Individual_Commodities).

This package should be coupled with the [GEKS package](https://github.com/MjStansfi/GEKS_package). Also see the alternates available for the FEWS (time product dummy) multilateral method [FEWS package](https://github.com/MjStansfi/FEWS_package) and
[TPD decomposition](https://github.com/MjStansfi/TPDDecomp).

See the attached pdf to understand how the code was simplified for conditions where either price a or b is missing.

# Installation


The GEKS decomposition function can be installed from GitHub using the following code
```{undefined eval=FALSE, include=TRUE}
devtools::install_github("MjStansfi/GEKSdecomp")
```

```{r}
# Once installed, the package can be loaded as usual
library(GEKSdecomp)
```

# Usage


The primary function provided by the the package is the `GEKS_decomp()` function. Running `?GEKS_decomp()` will give all the required information on how to use the function. An example of running the `GEKS_decomp()` the function is shown below.


# Example


This pacakge includes the Turvey dataset as found in the [Consumer Price Index Manual](https://www.ilo.org/wcmsp5/groups/public/---dgreports/---stat/documents/presentation/wcms_331153.pdf).

```{r, turvey-overview}
library(ggplot2)
ggplot(turvey, aes(x = month, y = price)) + 
  geom_line(aes(color = commodity)) + 
  geom_point(aes(color = commodity))+
  theme_bw() +
  ggtitle("Artificial prices of seasonal products Data created by R. Turvey")+
  annotate("rect", xmin=as.Date("1973-04-30"), xmax=as.Date("1973-05-31"), ymin=0, ymax=10, 
           alpha = .5)

```

The decomposition of the Turvey dataset with the contributions between 1973-04-30 and 1973-05-31 can be seen below.

``` {r}
#Make sure you restart your session if just installed

str(turvey)
# 'data.frame':	176 obs. of  4 variables:
#  $ month    : Date, format: "1970-01-31" "1970-01-31" "1970-01-31" "1970-02-28" ...
#  $ commodity: Factor w/ 5 levels "Apples","Grapes",..: 1 2 3 1 2 3 1 2 3 1 ...
#  $ price    : num  1.14 2.48 1.3 1.17 2.75 1.25 1.17 5.07 1.21 1.4 ...
#  $ quantity : int  3086 82 10266 3765 35 9656 4363 9 7940 4842 ...

contributions <- with(
    turvey,
    GEKS_decomp(times = month,
               price = price,
               id = commodity,
               weight = price*quantity,
               custom_time = c("1973-04-30","1973-05-31"),
               window_length = NULL)
 )

print(contributions)

```




```{r percentage_contrib, echo=FALSE}
ggplot(contributions)+
  geom_bar(aes(x=id,y=p_contrib), stat = "identity")+
  coord_flip()+
  ggtitle("Percentage contribution of observation")+
  geom_label(aes(x=id,y=p_contrib,label=paste0(round(p_contrib*100,1),"%")))+
  xlab("Product")+ylab("Percentage contribution")


```
